---
title: "Advertising"
author: "Griffin Kaiga"
date: "10/30/2020"
output: html_document
---

## 1.Defining the question 
A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process.
She wants to identify which individuals are most likely to click ads.

## 2. Metrics of sucess
The project will be considered a success once we are able to provide insights on the group of people who are most likely to click ads 

## 3. Understanding the context
Targeted advertising is a form of advertising, including online advertising, that is directed towards an audience with certain traits, based on the product or person the advertiser is promoting.These traits can either be demographic with a focus on race, economic status, sex, age, generation, level of education, income level, and employment, or there can be a psychographic focus which is based on the consumer values, personality, attitude, opinion, lifestyle and interest. This focus can also entail behavioral variables, such as browser history, purchase history, and other recent online activities. Targeted advertising is focused on certain traits and consumers who are likely to have a strong preference. These individuals will receive messages instead of those who have no interest and whose preferences do not match a particular product's attributes. This eliminates waste.

## 4. Experimental design
loading the data
checking the data
Find and deal with outliers, anomalies, and missing data within the dataset.
Perform  univariate and bivariate analysis

```{r}
# Viewing the datset
View(advertising)
class(advertising)
dim(advertising)

# checking the first six observations
head(advertising)

# checking the data types
str(advertising)
```
```{r}
# checking for the sum of missing values 
missing <- is.na(advertising)
colSums(missing)

# checking for outliers values
boxplot(advertising$Age)

# checking for the unique items
unique_items <- advertising[!duplicated(advertising),]
unique_items
```
# There were no missing values in the data.
# As for the boxplot for the column age, majority of the ages were between 30 and 44. The outliers valid since age is can extend to 60 years.
```{r}
# selecting the columns and making them all uniform
colnames(advertising)
names(advertising)[names(advertising) == 'Daily Time Spent on Site'] <- "time_spent"
names(advertising)[names(advertising) == "Daily Internet Usage"] <- "internet_usage"
names(advertising)[names(advertising) == "Clicked on Ad"] <- "clicked_on_ad" 
names(advertising)[names(advertising) == "Ad Topic Line"] <- "ad_topic_line"
names(advertising)[names(advertising) == "Area Income"] <- "area_income"
```
```{r}

# Univariate Graphical Exploratory Data Analysis

## 1. Measures of Central Tendency


# getting the summaries(min, max, 1st and 3rd quantiles, median, and mean)
summary(advertising)

# getting the mode
# creating a function that checks for the mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
   
# applying the function to the columns
# age
advertising.Age.mode <- getmode(advertising$Age)
advertising.Age.mode
# the most common age was 31 years

# time spent
advertising.time_spent.mode <- getmode(advertising$time_spent)
advertising.time_spent.mode
# the most common time spent was 62.26

# area of income
advertising.area_income.mode <- getmode(advertising$area_income)
advertising.area_income.mode
# the area income obtained often times was 61833.9

# internet usage
advertising.internet_usage.mode <- getmode(advertising$internet_usage)
advertising.internet_usage.mode
# the most appearing internet usage amount was 167.22

advertising.Country.mode <- getmode(advertising$Country)
advertising.Country.mode
#  "Czech Republic", was the frequent country in the blog

# city
advertising.City.mode <- getmode(advertising$City)
advertising.City.mode
# "Lisamouth",was the frequent city

# ad topic line
advertising.ad_topic_line.mode <- getmode(advertising$ad_topic_line)
advertising.ad_topic_line.mode
# the frequent ad topic line was, "Cloned 5thgeneration orchestration".
```

```{r}
# getting the range
advertising.time_spent.range <- range(advertising$time_spent)
advertising.time_spent.range

advertising.Age.range <- range(advertising$Age)
advertising.Age.range

advertising.area_income.range <- range(advertising$area_income)
advertising.area_income.range

advertising.internet_usage.range <- range(advertising$internet_usage)
advertising.internet_usage.range
```
# the range for the columns time spent, age, area income, and internet usage were 32.60 91.43
19 61, 13996.5 79484.8, and 104.78 269.96 respectively


```{r}
# getting the standard deviation
advertising.time_spent.sd <- sd(advertising$time_spent)
advertising.time_spent.sd

advertising.Age.sd <- sd(advertising$Age)
advertising.Age.sd

advertising.area_income.sd <- sd(advertising$area_income)
advertising.area_income.sd

advertising.internet_usage.sd <- sd(advertising$internet_usage)
advertising.internet_usage.sd
```
# the time spent, age, area_income, and internet usage had a standard deviation,  15.85361, 8.785562, 13414.63, and 43.90234 respectively


```{r}
# histograms for the numeric columns
hist(advertising$Age)

hist(advertising$time_spent)

hist(advertising$area_income)

hist(advertising$internet_usage)

hist(advertising$Male)
```
# the histograms indicate that the frequent ages are between 25years to 40 years. The least age is between 55-60 and 10-20

# The frequent time spent was between 65 to 85. The least appearing amount of time spent on the blog was between 30-35 and 90-95 

# the frequent chunk of income was between 50,000 to 70,000

# internet usage was frequent around 110 and 140, 200 to 240

# frequency for the column male(gender) was fairly equal


```{r}
## Bivariate Analysis

# determining the covariance
time_spent <- advertising$time_spent
age <- advertising$Age
click_on_ad <- advertising$clicked_on_ad
gender <- advertising$Male
internet_usage <- advertising$internet_usage
area_income <- advertising$area_income

cov(time_spent, age)
cov(click_on_ad, age)
cov(click_on_ad, gender)
cov(click_on_ad, time_spent)
cov(click_on_ad, internet_usage)
```
# The covariance of time spent and age duration is about -46.17415. It indicates a negative linear relationship between the two variables

# The covariance of time spent amd age duration is about 2.164665. It indicates a positive linear relationship between the two variables

# The covariance of time click on ad and gender is about -0.00950951. It indicates a zero since the click per ad is paired equally with both greater and lesser values of the gender varaible.

# The covariance of click on ad and time spent is about -5.933143. It indicates a negative linear relationship between the two variables

# The covariance of click on ad and internet usage is about  -17.27409. It indicates a negative linear relationship between the two variables


```{r}

# getting the correlation coefficients of the columns
cor(time_spent, age)
cor(click_on_ad, age)
cor(click_on_ad, gender)
cor(click_on_ad, time_spent)
cor(click_on_ad, internet_usage)
```
# the variables time spent and age have a weak negative linear correlation of -0.3315133

# the variables click on ad and age have a weak postive correlation of 0.4925313

# the variables "click on ad" and gender have an almost linear negative correlation of -0.7481166

# the variables "click on ad" and "internet usage have an almost linear negative correlation of -0.7865392  



## Recomendations 
Emphasis should be on the ages of between 35-45 since they are the majority in the blog.This is evident by the covariant relationship between "click on ad" and age variables.

Squeezing the cryptography course advertisment in the "Cloned 5thgeneration orchestration" ad topic would prove benefitial as evident by the mode of the ad topic line

More emphasis should be put on the country Czech Republic and the city "Lisamouth" as evident by the mode values of both
 




